---
title: "Frequently Asked Questions"
output:
  html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Frequently Asked Questions}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, message=FALSE, echo=FALSE}
library(unittest)
```

## General


### How do I test error conditions?

Use the ``ut_cmp_error`` function.
For example, here is a function that will throw an error for a bad argument:

```{r}
add_four <- function(x) {
    if( ! is.numeric(x) ) stop("x must be numeric")
    return( x+4 )
}
```

We can test the argument check like this:

```{r}
ok(ut_cmp_error(add_four("a"), "must be numeric"), "add_four() argument not numeric throws error")
```


### How do I test multivalue results?

Use ``all.equal(...)``:

```{r}
a <- c(1,2,3)
b <- 1:3
ok(all.equal(a,b), "a and b are equal")
```

Alternatively, the following helper function will give coloured output showing what's different

```{r}
cmp <- function(a, b) {
    if(identical(all.equal(a,b), TRUE)) return(TRUE)

    if (file.exists(Sys.which('git'))) {
        totmp <- function(x) {
            f <- tempfile(pattern = "str.")
            capture.output(str(x,
                vec.len = 1000,
                digits.d = 5,
                nchar.max = 1000), file = f)
            return(f)
        }

        return(suppressWarnings(system2(
            Sys.which('git'),
            c("diff", "--no-index", "--color-words", totmp(a), totmp(b)),
            input = "",
            stdout = TRUE, stderr = TRUE)))
    }

    return(c(
        capture.output(str(a)),
        "... does not equal...",
        capture.output(str(b))
    ))
}
```

To see the coloured output, try the following:

```{r, eval=FALSE}
ok(cmp(1:3, 1:8))
```


### Grouping tests

When dealing with many unit tests in one file it can be useful to group related unit tests.

The ``ok_group()`` function is used like this:

```{r}
ok_group("Test addition", {
    ok(1 + 1 == 2, "Can add 1")
    ok(1 + 3 == 4, "Can add 3")
})
ok_group("Test subtraction", {
    ok(1 - 1 == 0, "Can subtract 1")
    ok(1 - 3 == -2, "Can subtract 3")
})
```


### I am sure I do not need to test my code. Is this true?

No. Sit down and have a cup of tea. Hopefully the feeling will go away.


## Working with packages


### I'm writing a package, how do I put tests in it?

Add the following line to the package ``DESCRIPTION`` file.

```
Suggests: unittest
```

Create a directory called ``tests`` in your package source, alongside your ``R`` directory.

Place your tests in a file with the extension ``.R`` and add the following lines to the top of the file (replacing ``mypackage`` with the name of your package).


```{r, eval = FALSE}
library(mypackage)
library(unittest, quietly = TRUE)
```

That's it; ``R CMD check`` will run the tests and fail if any of the tests fail.

*Any* ``.R`` file in the ``tests`` directory will be run by ``R CMD check``.

When you use the ``unittest`` package the package 'knows' that it is being run by ``CMD check`` and at the end of the tests it produces a summary of the results.
The package will also throw an error if any tests fail;
throwing an error will in turn cause ``CMD check`` to report the error and fail the ``check``.

Here is a very simple example:

Assuming your package contains (and ``exports``) the function ``biggest()``

```{r}
biggest <- function(x,y) {max(c(x,y))}
```

then the ``tests/my_tests_for_biggest.R`` file could contain something like

```{r, eval = FALSE}
library(mypackage)
library(unittest, quietly = TRUE)

ok(biggest(3,4) == 4, "two numbers")
ok(biggest(c(5,3),c(3,4)) == 5, "two vectors")
```

### How do I test un-exported package functions?

If you have some unit tests which require access to un-exported functions,
or un-exported S3 methods, you can use ``local``.

```{r, eval = FALSE}
local({
    ok(internal_function() == 3)
    ok(another_internal_function() == 4)
    ok(final_internal_function() == 5)
}, asNamespace('mypackage'))
```
