---
title: "Frequently Asked Questions"
output:
  html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Frequently Asked Questions}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, message=FALSE, echo=FALSE}
library(unittest)
```

## General


### How do I test error conditions?

Use the ``ut_cmp_error`` function.
For example, here is a function that will throw an error for a bad argument:

```{r}
add_four <- function(x) {
    if( ! is.numeric(x) ) stop("x must be numeric")
    return( x+4 )
}
```

We can test the argument check like this:

```{r}
ok(ut_cmp_error(add_four("a"), "must be numeric"), "add_four() argument not numeric throws error")
```


### How do I test multivalue results, or see differences?

Use ``ut_cmp_equal(...)`` or ``ut_cmp_identical(...)`` as replacements for ``all.equal(...)`` and ``identical(...)`` respectively:

```{r}
a <- c(1,2,3)
b <- 1:3
ok(ut_cmp_equal(a,b), "a and b are equal")
```

``ut_cmp_identical`` will make sure your objects are identical, and is more
useful when comparing e.g. a list of strings which should be exactly the same.

``ut_cmp_equal`` will test for ‘near equality’, and is more useful when
comparing numeric values which may be slightly different due to floating-point
accuracy.

Either way, if your test fails you will get verbose output showing you how they
differ, and if you have git installed the output will be coloured. For example:

<pre>
> ok(ut_cmp_equal(c(1,2,3,4,5), c(1,8,8,4,5)))
not ok - ut_cmp_equal(c(1, 2, 3, 4, 5), c(1, 8, 8, 4, 5))
# Test returned non-TRUE value:
# Mean relative difference: 2.2
# --- c(1, 2, 3, 4, 5)
# +++ c(1, 8, 8, 4, 5)
# [1] 1 <span style="color:red;font-weight:bold">[-2 3-]</span><span style="color:green;font-weight:bold">{+8 8+}</span> 4 5
</pre>

### Grouping tests

When dealing with many unit tests in one file it can be useful to group related unit tests.

The ``ok_group()`` function is used like this:

```{r}
ok_group("Test addition", {
    ok(1 + 1 == 2, "Can add 1")
    ok(1 + 3 == 4, "Can add 3")
})
ok_group("Test subtraction", {
    ok(1 - 1 == 0, "Can subtract 1")
    ok(1 - 3 == -2, "Can subtract 3")
})
```


### I am sure I do not need to test my code. Is this true?

No. Sit down and have a cup of tea. Hopefully the feeling will go away.


## Working with packages


### I'm writing a package, how do I put tests in it?

Add the following line to the package ``DESCRIPTION`` file.

```
Suggests: unittest
```

Create a directory called ``tests`` in your package source, alongside your ``R`` directory.

Place your tests in a file with the extension ``.R`` and add the following lines to the top of the file (replacing ``mypackage`` with the name of your package).


```{r, eval = FALSE}
library(mypackage)
library(unittest, quietly = TRUE)
```

That's it; ``R CMD check`` will run the tests and fail if any of the tests fail.

*Any* ``.R`` file in the ``tests`` directory will be run by ``R CMD check``.

When you use the ``unittest`` package the package 'knows' that it is being run by ``CMD check`` and at the end of the tests it produces a summary of the results.
The package will also throw an error if any tests fail;
throwing an error will in turn cause ``CMD check`` to report the error and fail the ``check``.

Here is a very simple example:

Assuming your package contains (and ``exports``) the function ``biggest()``

```{r}
biggest <- function(x,y) {max(c(x,y))}
```

then the ``tests/my_tests_for_biggest.R`` file could contain something like

```{r, eval = FALSE}
library(mypackage)
library(unittest, quietly = TRUE)

ok(biggest(3,4) == 4, "two numbers")
ok(biggest(c(5,3),c(3,4)) == 5, "two vectors")
```

### How do I test un-exported package functions?

If you have some unit tests which require access to un-exported functions,
or un-exported S3 methods, you can use ``local``.

```{r, eval = FALSE}
local({
    ok(internal_function() == 3)
    ok(another_internal_function() == 4)
    ok(final_internal_function() == 5)
}, asNamespace('mypackage'))
```
