---
title: "Getting Started"
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## A very simple example

Suppose you have a file ``biggest.R`` with the following function:

```{r}
biggest <- function(x,y) {max(c(x,y))}
```

To test this create a file called ``test_biggest.R`` in the same directory containing:

```{r, eval=FALSE}
library(unittest, quietly = TRUE)

source('biggest.R')

ok(biggest(3,4) == 4, "two numbers")
ok(biggest(c(5,3),c(3,4)) == 5, "two vectors")
```

Now in an ``R`` session ``source`` the test file:

```{r, eval=FALSE}
source('test_biggest.R')
```

and you will see output like this

```
ok - two numbers
ok - two vectors
```

and that's it.

Now each time you edit ``biggest.R`` re-sourcing ``test_biggest.R``
reloads your function and runs your unit tests.

## Comparing results

Suppose our ``biggest`` function was broken, for example:

```{r}
biggest <- function(x,y) { 4 }
```

Our tests from earlier would return:

```
ok - two numbers
not ok - two vectors
# Test returned non-TRUE value:
# [1] FALSE
```

It would be useful to see what did get returned, to help work out the problem.

To help with this we can use ``ut_cmp_equal``. If we rewrite our test to:

```{r, eval=FALSE}
library(unittest, quietly = TRUE)

source('biggest.R')

ok(ut_cmp_equal(biggest(3,4), 4), "two numbers")
ok(ut_cmp_equal(biggest(c(5,3),c(3,4)), 5), "two vectors")
```

Now the test output will show both what we got and what the test says we should get:

```
ok - two numbers
not ok - two vectors
# Test returned non-TRUE value:
# Mean relative difference: 0.25
# --- biggest(c(5, 3), c(3, 4))
# +++ 5
# [1] [-4-]{+5+}
```
